#Reproducible Research Project 1
###By Andy Ng

This rmarkdown file is used to answer the question in project 1
First setup the R envirnoment
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Here we load the data into raw data
```{r load data}
  #Load data
  raw_steps_data <- read.csv("activity.csv")
```

###Here we remove the NA and calculate the total steps per day, mean per day, and median per day.
```{r no. of steps}
  
  #Remove NA's
  steps_data <- na.omit(raw_steps_data)
  #SUM data per date
  steps_data <- aggregate(steps_data$steps, by= list(steps_data$date), FUN = sum)
  #Set column names
  colnames(steps_data) <- c("Dates", "Steps")
  #Convert date object
  steps_data[,1] <- as.Date(as.character(steps_data[,1]), format = "%Y-%m-%d")
  #Calculation
  mean_per_day <- mean(as.numeric(steps_data$Steps))
  median_per_day <- median(as.numeric(steps_data$Steps))
  #plotting the no. of steps per day
  barplot(steps_data$Steps, names = steps_data$Date, xlab = "Date", ylab = "No. of Steps", main = "Number of steps per day")
```


The mean per day is `r as.character(mean_per_day)`

The mean per day is `r as.character(median_per_day)`

###Here we calculate the interval data
```{r 5 minutes interval plot}
  # remove NA's from the data
  interval_steps_data <- na.omit(raw_steps_data)
  # calculate the mean per interval
  interval_steps_data <- aggregate(interval_steps_data$steps, by=list(interval_steps_data$interval), FUN = mean)
  # set the column names
  colnames(interval_steps_data) <- c("Interval", "Steps")
  # plot the data
  plot(interval_steps_data$Interval, interval_steps_data$Steps, xlab = "Interval", ylab = "No. of steps", main = "Number of steps by interval across all days", type = "l")
  max_interval_data <- subset(interval_steps_data, Steps == max(interval_steps_data$Steps))
```

The interval that contains the maximum number of steps is `r max_interval_data[1,1]`

###Here we provide a solution for the NA's. We will use the average of that interval across all data to fill in the NA's

The number of NA rows is `r nrow(raw_steps_data) - nrow(na.omit(raw_steps_data))`

```{r replace missing value}
  ##Load data
  replaced_data <- raw_steps_data
  ##calculate the mean interval data
  mean_interval_data <- na.omit(raw_steps_data)
  mean_interval_data <- aggregate(mean_interval_data$steps, by = list(mean_interval_data$interval), FUN = mean)
  colnames(mean_interval_data) <- c("Interval", "Steps")
  ##replace the NA's with mean interval data
  replaced_data$steps <- ifelse(is.na(replaced_data$steps), mean_interval_data$Steps[mean_interval_data$Interval %in% replaced_data$interval], replaced_data$steps)
  ## calcuate the sum of data
  replaced_data <- aggregate(replaced_data$steps, by = list(replaced_data$date), FUN = sum)
  colnames(replaced_data) <- c("Date", "Steps")
  ##Convert date object
  replaced_data[,1] <- as.Date(as.character(replaced_data[,1]), format = "%Y-%m-%d")
  ## plot the data
  barplot(replaced_data$Steps, names = replaced_data$Date)
```

The mean of the replaced data is `r as.character(mean(as.numeric(replaced_data$Steps)))`

The median of the replaced data is `r as.character(median(as.numeric(replaced_data$Steps)))`

In conclusion, after we filled in all the NA's. The mean remains the same and the median has become the same as the mean.

###Here we analysis the difference between weekday and weekend and provide a plot for it

```{r weekday weekend analysis}
  ##Get raw data
  week_data <- na.omit(raw_steps_data)
  ##Convert date object
  week_data[,2] <- as.Date(as.character(week_data[,2]), format = "%Y-%m-%d")
  ##add a column to determine weekday adn weekend
  week_data[,4] <- ifelse((as.POSIXlt(week_data[,2])$wday == 0 | as.POSIXlt(week_data[,2])$wday == 6), "Weekend", "Weekday")
  ##separate weekday and weekend data
  weekday_data <- week_data[week_data[,4] == "Weekday",]
  weekend_data <- week_data[week_data[,4] == "Weekend",]
  ## calculate the data
  weekday_data <- aggregate(weekday_data$steps, by = list(weekday_data$interval), FUN = mean)
  weekend_data <- aggregate(weekend_data$steps, by = list(weekend_data$interval), FUN = mean)
  colnames(weekday_data) <- c("Interval", "Steps")
  colnames(weekend_data) <- c("Interval", "Steps")
  ##Plot the data
  par(mfrow = c(2,1), mar = c(2,2,2,2))
  plot(weekday_data$Interval, weekday_data$Steps, type = "l", xlab = "Interval", ylab = "No. of steps", main = "Weekday Data")
  plot(weekend_data$Interval, weekend_data$Steps, type = "l", xlab = "Interval", ylab = "No. of steps", main = "Weekend Data")
```